Decisions Decisions Decisions: Carrying out Latent Profile Analysis in Accordance With Best Practices Using the tidyLPA R package

### 2020-12-28 paper.rmd
update

```{r setup, include = FALSE}
library(papaja)
```

```{r analysis-preferences, echo = FALSE}
# Seed for random number generation
set.seed(42)
library(tidyverse)
library(simstudy)

@@

@@

```{r}
pisaUSA15 %>%
  select(broad_interest, enjoyment, self_efficacy) %>% 
  psych::describe()
```

@@

``` {r, curry-mac-mclust-example}
pisaUSA15 %>%
  select(broad_interest, enjoyment, self_efficacy) %>%
  single_imputation() %>%
  estimate_profiles(3)
```

@@

``` {r, curry-mac-mplus, eval = FALSE}
pisaUSA15 %>%
  select(broad_interest, enjoyment, self_efficacy) %>%
  single_imputation() %>%
  estimate_profiles(3, package = "mplus")
```

@@

```{r, get-fit-m1}
pisaUSA15 %>%
  select(broad_interest, enjoyment, self_efficacy) %>%
  single_imputation() %>%
  estimate_profiles(3) %>% 
  get_fit(m1) %>% 
  knitr::kable()
```

@@

````{r, compare-mclust, eval= TRUE}
compare_solutions_results <- pisaUSA15 %>%
  select(broad_interest, enjoyment, self_efficacy) %>%
  single_imputation() %>%
  estimate_profiles(models = c(1, 2, 3, 6),
                    n_profiles = 1:9)

compare_solutions_results

compare_solutions(compare_solutions_results)
```

@@

# Recommendations Regarding Analytic Choices

>>>>

```{r setup, include = FALSE}
library(papaja)
# Seed for random number generation
set.seed(42)
library(tidyverse)
library(simstudy)

@@

## Recommendations Regarding Analytic Choices

@@

```{r}
pisaUSA15 %>%
  select(broad_interest, enjoyment, self_efficacy) %>% 
  psych::describe() %>% 
  knitr::kable()
```

@@

``` {r, curry-mac-mclust-example, echo = TRUE}
pisaUSA15 %>%
  select(broad_interest, enjoyment, self_efficacy) %>%
  single_imputation() %>%
  estimate_profiles(3)
```

@@

``` {r, curry-mac-mplus, echo = TRUE}
pisaUSA15 %>%
  select(broad_interest, enjoyment, self_efficacy) %>%
  single_imputation() %>%
  estimate_profiles(3, package = "mplus")
```

@@

```{r, get-fit-m1}
pisaUSA15 %>%
  select(broad_interest, enjoyment, self_efficacy) %>%
  single_imputation() %>%
  estimate_profiles(3) %>% 
  get_fit(m1)
```

@@

````{r, compare-mclust, eval= TRUE}
many_analyses <- pisaUSA15 %>%
  select(broad_interest, enjoyment, self_efficacy) %>%
  single_imputation() %>%
  estimate_profiles(models = 1,
                    n_profiles = 1:6)

many_analyses

compare_solutions(many_analyses)

many_analyses_1_3 <- pisaUSA15 %>%
  select(broad_interest, enjoyment, self_efficacy) %>%
  single_imputation() %>%
  estimate_profiles(models = c(1, 3),
                    n_profiles = 1:6)

many_analyses_1_3

compare_solutions(many_analyses_1_3)
```

It appears that the four and five-class solutions fit best. Let's inspect those, specifically:

```{r}
models_4_5 <- pisaUSA15[1:100, ] %>%
  select(broad_interest, enjoyment, self_efficacy) %>%
  single_imputation() %>%
  estimate_profiles(models = 1,
                    n_profiles = c(4, 5))

models_4_5 %>% 
  plot_profiles(add_line = TRUE)
```

We can also create a faceted plot of density plots for an object of class 'tidyLPA'. For each variable, a Total density plot will be shown, along with separate density plots for each latent class, where cases are weighted by the posterior probability of being assigned to that class:

```{r}
models_4_5 %>% 
  plot_profiles(add_line = TRUE)
```

Last, we can create a faceted plot of two-dimensional correlation plots and unidimensional density plots for an object of class 'tidyProfile':

```{r}
models_4_5 %>% 
  plot_bivariate()
```

```{r, eval = FALSE, echo = FALSE}
many_analyses_mplus <- pisaUSA15 %>%
  select(broad_interest, enjoyment, self_efficacy) %>%
  single_imputation() %>%
  estimate_profiles(models = c(1),
                    n_profiles = 1:7,
                    package = "mplus")

many_analyses_mplus

compare_solutions(many_analyses_mplus)
```

@@

### 2020-12-28 paper.tex
update

\PassOptionsToPackage{unicode=true}{hyperref} % options for packages loaded elsewhere
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[man]{apa6}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provides euro and other symbols
\else % if luatex or xelatex
  \usepackage{unicode-math}
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage[]{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\usepackage{hyperref}
\hypersetup{
            pdftitle={Decisions Decisions Decisions: Carrying out Latent Profile Analysis in Accordance With Best Practices Using the tidyLPA R package},
            pdfauthor={Joshua Rosenberg, Caspar van Lissa, Jennifer Schmidt, Patrick Beymer, Daniel Anderson, \& Matthew Schell},
            pdfkeywords={Latent Profile Analysis, mixture models, finite mixture models, tutorial, R, MPlus, mclust},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

@@

\begin{verbatim}
## Warning: package 'ggplot2' was built under R version 3.6.2
\end{verbatim}

\begin{verbatim}
## Warning: package 'tibble' was built under R version 3.6.2
\end{verbatim}

\begin{verbatim}
## Warning: package 'tidyr' was built under R version 3.6.2
\end{verbatim}

\begin{verbatim}
## Warning: package 'readr' was built under R version 3.6.2
\end{verbatim}

\begin{verbatim}
## Warning: package 'purrr' was built under R version 3.6.2
\end{verbatim}

\begin{verbatim}
## Warning: package 'dplyr' was built under R version 3.6.2
\end{verbatim}

\begin{verbatim}
## Warning: package 'simstudy' was built under R version 3.6.2
\end{verbatim}

\begin{verbatim}
## NULL
\end{verbatim}

@@

\hypertarget{specific-software-tools}{%
\section{Specific Software Tools}\label{specific-software-tools}}

@@

\hypertarget{tidylpa}{%
\subsection{tidyLPA}\label{tidylpa}}

@@

Here, we use a simulated dataset contains data on moral relevance and judgment across the seven domains of the \emph{Morality As Cooperation} scale. More on this scale can be found in Curry et al., (2019).

@@

@@

\begin{tabular}{r|r|r|r|r|r|r|r|r|r|r|r|r|r|r|r|r|r}
\hline
Model & Classes & LogLik & AIC & AWE & BIC & CAIC & CLC & KIC & SABIC & ICL & Entropy & prob\_min & prob\_max & n\_min & n\_max & BLRT\_val & BLRT\_p\\
\hline
1 & 1 & -81245.00 & 162546.0 & 162977.4 & 162692.7 & 162720.7 & 162492.0 & 162577.0 & 162603.7 & -162692.7 & 1.0000000 & 1.0000000 & 1.0000000 & 1.0000000 & 1.0000000 & NA & NA\\
\hline
1 & 2 & -79727.50 & 159541.0 & 160204.8 & 159766.3 & 159809.3 & 159456.7 & 159587.0 & 159629.7 & -159925.3 & 0.8369494 & 0.9296991 & 0.9637189 & 0.3441092 & 0.6558908 & 3035.00346 & 0.009901\\
\hline
1 & 3 & -79239.27 & 158594.5 & 159490.6 & 158898.4 & 158956.4 & 158480.2 & 158655.5 & 158714.1 & -159175.7 & 0.8188766 & 0.9053581 & 0.9222054 & 0.1465517 & 0.5272989 & 976.44661 & 0.009901\\
\hline
1 & 4 & -79090.59 & 158327.2 & 159455.5 & 158709.6 & 158782.6 & 158182.7 & 158403.2 & 158477.7 & -159150.2 & 0.7763719 & 0.8393595 & 0.9127306 & 0.0732759 & 0.4655172 & 297.35974 & 0.009901\\
\hline
1 & 5 & -79029.50 & 158235.0 & 159595.6 & 158696.0 & 158784.0 & 158060.4 & 158326.0 & 158416.4 & -159398.5 & 0.7012112 & 0.6979575 & 0.8987374 & 0.0675287 & 0.2406609 & 122.19715 & 0.009901\\
\hline
1 & 6 & -78971.30 & 158148.6 & 159741.3 & 158688.2 & 158791.2 & 157944.0 & 158254.6 & 158361.0 & -159456.7 & 0.6954594 & 0.6929238 & 0.9230731 & 0.0603448 & 0.3060345 & 116.39215 & 0.009901\\
\hline
1 & 7 & -78930.08 & 158096.2 & 159921.0 & 158714.3 & 158832.3 & 157861.5 & 158217.2 & 158339.5 & -159550.6 & 0.6933730 & 0.5964096 & 0.9236077 & 0.0610632 & 0.2550287 & 82.44963 & 0.009901\\
\hline
1 & 8 & -78851.80 & 157969.6 & 160026.5 & 158666.3 & 158799.3 & 157705.1 & 158105.6 & 158243.8 & -159391.4 & 0.7477064 & 0.6432919 & 0.9219048 & 0.0423851 & 0.2902299 & 156.55019 & 0.009901\\
\hline
1 & 9 & -78821.74 & 157939.5 & 160228.6 & 158714.8 & 158862.8 & 157645.0 & 158090.5 & 158244.6 & -159455.8 & 0.7515427 & 0.6071032 & 0.9456474 & 0.0402299 & 0.2938218 & 60.12578 & 0.009901\\
\hline
1 & 10 & -78751.27 & 157828.5 & 160349.8 & 158682.4 & 158845.4 & 157504.1 & 157994.5 & 158164.6 & -159420.9 & 0.7625067 & 0.6463008 & 0.9396876 & 0.0337644 & 0.2909483 & 140.92881 & 0.009901\\
\hline
\end{tabular}

@@

Here is an example of specifying a model with varying variances and covariances
(Model 6)

\begin{verbatim}
## tidyLPA analysis using mclust: 
## 
##  Model Classes AIC       BIC       Entropy prob_min prob_max n_min n_max BLRT_p
##  6     3       156872.22 158752.84 0.67    0.82     0.87     0.23  0.41  0.01
\end{verbatim}

\begin{verbatim}
## tidyLPA analysis using mclust: 
## 
##  Model Classes AIC       BIC       Entropy prob_min prob_max n_min n_max BLRT_p
##  1     1       139226.33 139352.05 1.00    1.00     1.00     1.00  1.00        
##  1     2       136511.15 136704.97 0.83    0.92     0.97     0.34  0.66  0.01  
##  1     3       135651.76 135913.69 0.81    0.90     0.92     0.14  0.53  0.01  
##  1     4       135391.00 135721.03 0.77    0.83     0.91     0.07  0.47  0.01  
##  1     5       135326.12 135724.24 0.72    0.57     0.91     0.07  0.39  0.01  
##  1     6       135281.64 135747.86 0.68    0.56     0.92     0.07  0.28  0.01  
##  1     7       135218.07 135752.39 0.67    0.59     0.91     0.07  0.29  0.01  
##  1     8       135135.62 135738.04 0.70    0.59     0.92     0.05  0.31  0.01  
##  1     9       135098.58 135769.10 0.70    0.57     0.93     0.05  0.30  0.01  
##  2     1       139226.33 139352.05 1.00    1.00     1.00     1.00  1.00        
##  2     2       136245.76 136502.44 0.81    0.92     0.96     0.39  0.61  0.01  
##  2     3       135337.03 135724.68 0.80    0.91     0.91     0.25  0.49  0.01  
##  2     4       135103.11 135621.72 0.75    0.82     0.92     0.17  0.37  0.01  
##  2     5       134840.49 135490.07 0.75    0.76     0.89     0.13  0.31  0.01  
##  2     6       134688.81 135469.35 0.75    0.71     0.89     0.06  0.29  0.01  
##  2     7       134579.47 135490.97 0.75    0.64     0.90     0.06  0.23  0.01  
##  2     8       134482.30 135524.76 0.75    0.71     0.90     0.06  0.18  0.01  
##  2     9       134434.76 135608.19 0.75    0.65     0.91     0.06  0.17  0.01  
##  3     1       134912.97 135384.44 1.00    1.00     1.00     1.00  1.00        
##  3     2       134836.15 135375.72 0.69    0.78     0.96     0.22  0.78  0.01  
##  3     3       134862.13 135469.80 0.41    0.16     0.88     0.07  0.69  0.76  
##  3     4       134728.94 135404.71 0.50    0.31     0.83     0.10  0.63  0.01  
##  3     5       134615.77 135359.63 0.58    0.43     0.79     0.07  0.56  0.01  
##  3     6       134696.35 135508.32 0.52    0.43     0.81     0.06  0.39  0.71  
##  3     7       134589.15 135469.21 0.54    0.36     0.84     0.06  0.37  0.01  
##  3     8       134604.12 135552.29 0.54    0.40     0.84     0.04  0.33  0.77  
##  3     9       134568.05 135584.32 0.56    0.31     0.83     0.03  0.34  0.01  
##  6     1       134912.97 135384.44 1.00    1.00     1.00     1.00  1.00        
##  6     2       134223.62 135171.78 0.62    0.87     0.90     0.48  0.52  0.01  
##  6     3       134151.67 135576.55 0.68    0.80     0.88     0.18  0.47  0.01  
##  6     4       134089.83 135991.41 0.68    0.77     0.87     0.17  0.36  0.01  
##  6     5       134092.59 136470.87 0.71    0.77     0.89     0.06  0.33  0.34  
##  6     6       134088.85 136943.83 0.72    0.75     0.90     0.06  0.29  0.43  
##  6     7       134021.25 137352.94 0.74    0.77     0.88     0.09  0.20  0.01  
##  6     8       134038.51 137846.90 0.77    0.79     0.98     0.03  0.19  0.58  
##  6     9       134067.02 138352.11 0.78    0.77     0.96     0.04  0.18  0.78
\end{verbatim}

\begin{verbatim}
## Compare tidyLPA solutions:
## 
##  Model Classes BIC       
##  1     1       139352.049
##  1     2       136704.970
##  1     3       135913.687
##  1     4       135721.026
##  1     5       135724.243
##  1     6       135747.862
##  1     7       135752.395
##  1     8       135738.044
##  1     9       135769.104
##  2     1       139352.049
##  2     2       136502.441
##  2     3       135724.680
##  2     4       135621.724
##  2     5       135490.065
##  2     6       135469.349
##  2     7       135490.971
##  2     8       135524.760
##  2     9       135608.188
##  3     1       135384.436
##  3     2       135375.718
##  3     3       135469.796
##  3     4       135404.708
##  3     5       135359.634
##  3     6       135508.318
##  3     7       135469.214
##  3     8       135552.290
##  3     9       135584.321
##  6     1       135384.436
##  6     2       135171.784
##  6     3       135576.546
##  6     4       135991.409
##  6     5       136470.871
##  6     6       136943.826
##  6     7       137352.936
##  6     8       137846.896
##  6     9       138352.109
## 
## Best model according to BIC is Model 6 with 2 classes.
## 
## An analytic hierarchy process, based on the fit indices AIC, AWE, BIC, CLC, and KIC (Akogul & Erisoglu, 2017), suggests the best solution is Model 6 with 2 classes.
\end{verbatim}

\includegraphics{paper_files/figure-latex/plot-one-solution-1.pdf}

@@

\hypertarget{recommendations-regarding-analytic-choices}{%
\section{Recommendations Regarding Analytic Choices}\label{recommendations-regarding-analytic-choices}}

\hypertarget{htmlwidget-675d126ab2afaf44720a}{}

\hypertarget{how-the-data-are-centering-or-scaled}{%
\subsection{How the Data are Centering or Scaled}\label{how-the-data-are-centering-or-scaled}}

The data can be transformed by centering or scaling. Typically, these are done
after the profiles are created, so that differences between profiles can be more
easily explored. They can also be done prior to the analysis, which can be
helpful for obtaining solutions when the variables are on very different
scales.

\hypertarget{how-to-choose-the-number-of-profilesclasses}{%
\subsection{How to Choose the Number of Profiles/Classes}\label{how-to-choose-the-number-of-profilesclasses}}

\hypertarget{how-to-choose-options-for-the-clustering-procedure-or-to-select-the-model-parameterization}{%
\subsection{How to Choose Options for the Clustering Procedure or to Select the Model Parameterization}\label{how-to-choose-options-for-the-clustering-procedure-or-to-select-the-model-parameterization}}

\hypertarget{more-general-recommendations}{%
\subsection{More General Recommendations}\label{more-general-recommendations}}

This section highlights three general recommendations. First, scholars taking a
person-oriented approach should emphasize reproducibility in carrying out
analyses. This is in part due to the exploratory and interpretative nature of
person-oriented approaches. To this end, presenting multiple models, as in
Pastor et al. (2007), should be encouraged, rather than presenting one solution.
In addition, having multiple analysts review the solutions found is encouraged.
As part of this recommendation, we suggest that researchers consider how their
analyses can be reproduced by other analysts outside of the research team:
Sharing code and data is an important part of this work. Also related,
researchers should consider how profiles are replicated across samples.

>>>>

\PassOptionsToPackage{unicode=true}{hyperref} % options for packages loaded elsewhere
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[man]{apa6}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provides euro and other symbols
\else % if luatex or xelatex
  \usepackage{unicode-math}
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage[]{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
\usepackage{hyperref}
\hypersetup{
            pdftitle={Decisions Decisions Decisions: Carrying out Latent Profile Analysis in Accordance With Best Practices Using the tidyLPA R package},
            pdfauthor={Joshua Rosenberg, Caspar van Lissa, Jennifer Schmidt, Patrick Beymer, Daniel Anderson, \& Matthew Schell},
            pdfkeywords={Latent Profile Analysis, mixture models, finite mixture models, tutorial, R, MPlus, mclust},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{248,248,248}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\BuiltInTok}[1]{#1}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
\newcommand{\ExtensionTok}[1]{#1}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ImportTok}[1]{#1}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\NormalTok}[1]{#1}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\RegionMarkerTok}[1]{#1}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

@@

@@

\hypertarget{best-practices-for-latent-profile-analysis}{%
\section{Best Practices for Latent Profile Analysis}\label{best-practices-for-latent-profile-analysis}}

\hypertarget{recommendations-regarding-analytic-choices}{%
\subsection{Recommendations Regarding Analytic Choices}\label{recommendations-regarding-analytic-choices}}

\hypertarget{htmlwidget-6725a3bb4d9d4a47b614}{}

\hypertarget{future-directions-and-extensions}{%
\section{Future Directions and Extensions}\label{future-directions-and-extensions}}

First, scholars taking a
person-oriented approach should emphasize reproducibility in carrying out
analyses. This is in part due to the exploratory and interpretative nature of
person-oriented approaches. To this end, presenting multiple models, as in
Pastor et al. (2007), should be encouraged, rather than presenting one solution.
In addition, having multiple analysts review the solutions found is encouraged.
As part of this recommendation, we suggest that researchers consider how their
analyses can be reproduced by other analysts outside of the research team:
Sharing code and data is an important part of this work. Also related,
researchers should consider how profiles are replicated across samples.

\hypertarget{tidylpa}{%
\section{tidyLPA}\label{tidylpa}}

@@

\hypertarget{the-tidylpa-r-package}{%
\subsection{The tidyLPA R package}\label{the-tidylpa-r-package}}

@@

Here, we use a simulated dataset contains data on broad interest, enjoyment, and self-efficacy from the PISA assessment.

First, let's explore the data graphically using the \{ggplot2\} package:

\includegraphics{paper_files/figure-latex/unnamed-chunk-7-1.pdf}

Here, we use the \{psych\} package to explore the distributional properties of the variables:

\begin{tabular}{l|r|r|r|r|r|r|r|r|r|r|r|r|r}
\hline
  & vars & n & mean & sd & median & trimmed & mad & min & max & range & skew & kurtosis & se\\
\hline
broad\_interest & 1 & 5454 & 2.652246 & 0.7772383 & 2.8 & 2.683902 & 0.59304 & 1 & 5 & 4 & -0.2767312 & -0.0261463 & 0.0105244\\
\hline
enjoyment & 2 & 5520 & 2.780263 & 0.7509386 & 3.0 & 2.819973 & 0.59304 & 1 & 4 & 3 & -0.5076839 & 0.1652595 & 0.0101073\\
\hline
self\_efficacy & 3 & 5409 & 2.136430 & 0.6729143 & 2.0 & 2.112135 & 0.63540 & 1 & 4 & 3 & 0.4376698 & 0.0705830 & 0.0091496\\
\hline
\end{tabular}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{pisaUSA15 }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{select}\NormalTok{(broad_interest, enjoyment, self_efficacy) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{single_imputation}\NormalTok{() }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{estimate_profiles}\NormalTok{(}\DecValTok{3}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## tidyLPA analysis using mclust: 
## 
##  Model Classes AIC      BIC      Entropy prob_min prob_max n_min n_max BLRT_p
##  1     3       34815.05 34908.15 0.75    0.77     0.94     0.16  0.67  0.01
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{pisaUSA15 }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{select}\NormalTok{(broad_interest, enjoyment, self_efficacy) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{single_imputation}\NormalTok{() }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{estimate_profiles}\NormalTok{(}\DecValTok{3}\NormalTok{, }\DataTypeTok{package =} \StringTok{"mplus"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## tidyLPA analysis using mplus: 
## 
##  Model Classes AIC      BIC      Entropy prob_min prob_max n_min n_max BLRT_p
##  1     3       34804.95 34898.05 0.77    0.79     0.94     0.16  0.67  0.00
\end{verbatim}

\hypertarget{fit}{%
\subsubsection{Fit}\label{fit}}

@@

\begin{verbatim}
## # A tibble: 1 x 18
##   Model Classes  LogLik    AIC    AWE    BIC   CAIC    CLC    KIC  SABIC     ICL
##   <dbl>   <dbl>   <dbl>  <dbl>  <dbl>  <dbl>  <dbl>  <dbl>  <dbl>  <dbl>   <dbl>
## 1     1       3 -17386. 34800. 35055. 34893. 34907. 34774. 34817. 34849. -36424.
## # ... with 7 more variables: Entropy <dbl>, prob_min <dbl>, prob_max <dbl>,
## #   n_min <dbl>, n_max <dbl>, BLRT_val <dbl>, BLRT_p <dbl>
\end{verbatim}

\begin{verbatim}
## tidyLPA analysis using mclust: 
## 
##  Model Classes AIC      BIC      Entropy prob_min prob_max n_min n_max BLRT_p
##  1     1       37962.61 38002.51 1.00    1.00     1.00     1.00  1.00        
##  1     2       35495.33 35561.83 0.74    0.83     0.96     0.24  0.76  0.01  
##  1     3       34850.99 34944.09 0.75    0.76     0.94     0.16  0.67  0.01  
##  1     4       32990.63 33110.33 0.93    0.94     0.97     0.07  0.54  0.01  
##  1     5       32652.78 32799.09 0.89    0.79     0.99     0.07  0.46  0.01  
##  1     6       32631.88 32804.79 0.76    0.44     0.99     0.07  0.34  0.01
\end{verbatim}

\begin{verbatim}
## Compare tidyLPA solutions:
## 
##  Model Classes BIC      
##  1     1       38002.509
##  1     2       35561.831
##  1     3       34944.094
##  1     4       33110.333
##  1     5       32799.090
##  1     6       32804.788
## 
## Best model according to BIC is Model 1 with 5 classes.
## 
## An analytic hierarchy process, based on the fit indices AIC, AWE, BIC, CLC, and KIC (Akogul & Erisoglu, 2017), suggests the best solution is Model 1 with 6 classes.
\end{verbatim}

@@

@@
